# Baseline model

The goal of this documents is to compare some tuned and untuned models on a "baseline" dataset. We want to predict whether AE in 2020 goes above 3. We are only using 2019 data here! Next steps will be adding time-dependent data (e.g. COVID deaths per week per zipcode).

## Common data
```{r}
library(tidyverse)
library(tidymodels)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
per <- read_feather("data/simulation_data/all_persons.feather")
```

Compute some summary statistic for each client.
```{r}
clients <-
  per %>%
  group_by(client) %>%
  summarize(
    zip3 = first(zip3),
    size = n(),
    volume = sum(FaceAmt),
    avg_qx = mean(qx),
    avg_age = mean(Age),
    per_male = sum(Sex == "Male") / size,
    per_blue_collar = sum(collar == "blue") / size,
    expected = sum(qx * FaceAmt),
    actual_2020 = sum(FaceAmt[year == 2020], na.rm = TRUE),
    ae_2020 = actual_2020 / expected,
    actual_2019 = sum(FaceAmt[year == 2019], na.rm = TRUE),
    ae_2019 = actual_2019 / expected,
    adverse = as_factor(if_else(ae_2020 > 3, "ae > 3", "ae < 3"))
  ) %>%
  relocate(adverse, ae_2020, .after = zip3) %>%
  mutate(adverse = fct_relevel(adverse, c("ae > 3", "ae < 3")))
```

We can add some demographic information based on zip3.
```{r}
zip_data <-
  read_feather("data/data.feather") %>%
  mutate(
    density = POP / AREALAND,
    AREALAND = NULL,
    AREA = NULL,
    HU = NULL,
    vaccinated = NULL,
    per_lib = NULL,
    per_green = NULL,
    per_other = NULL,
    per_rep = NULL,
    unempl_2020 = NULL,
    deaths_covid = NULL,
    deaths_all = NULL
  ) %>%
  rename(
    unemp = unempl_2019,
    hes_uns = hes_unsure,
    str_hes = strong_hes,
    income = Median_Household_Income_2019
  )
```
There seems to be some clients with some zip codes that we cannot deal with. These are the ones
```{r}
clients %>%
  anti_join(zip_data, by = "zip3") %>%
  select(zip3)
```
These correspond to the following areas

ZIP3 | Area       |
-----|------------|
969  | Guam, Palau, Federated States of Micronesia, Northern Mariana Islands, Marshall Islands |
093  | Military bases in Iraq and Afghanistan |
732  | Not in use |
872  | Not in use |
004  | Not in use |
202  | Washington DC, Government 1 |

We ignore clients with these zip codes. There are also two clients in DC for which we're missing election data. We will ignore those as well.
```{r}
clients %<>%
  inner_join(zip_data, by = "zip3") %>%
  drop_na()
```

We now have our full dataset. Behold!
```{r}
skim(clients)
```

## Workflow set
We'll evaluate models using a workflow set. To make our life easier, we will remove some variables and use a formula instead of a recipe.
```{r}
clients <-
  clients %>%
  select(-client, -zip3, -ae_2020, -actual_2020, -actual_2019)
```

We now gather our recipes and models.
```{r}
with2019_rec <-
  recipe(adverse ~ ., data = clients) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors(), -all_nominal())
no2019_rec <-
  with2019_rec %>%
  step_rm(ae_2019)

log_spec <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
tuned_log_spec <-
  logistic_reg(penalty = 0.00118) %>%
  set_engine("glmnet") %>%
  set_mode("classification")
forest_spec <-
  rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
tuned_forest_spec <-
  rand_forest(trees = 1000, mtry = 12, min_n = 21) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
sln_spec <-
  mlp() %>%
  set_engine("nnet") %>%
  set_mode("classification")
svm_rbf_spec <-
  svm_rbf() %>%
  set_engine("kernlab") %>%
  set_mode("classification")
svm_poly_spec <-
  svm_poly() %>%
  set_engine("kernlab") %>%
  set_mode("classification")
knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")

models <- list(log = log_spec,
               logtuned = tuned_log_spec,
               forest = forest_spec,
               foresttuned = tuned_forest_spec,
               neural = sln_spec,
               svmrbf = svm_rbf_spec,
               svmpoly = svm_poly_spec,
               knnspec = knn_spec)
recipes <- list("with2019ae" = with2019_rec,
                "no2019ae" = no2019_rec)
wflows <- workflow_set(recipes, models)
```

Data splitting
```{r}
set.seed(30308)
init <- initial_split(clients, strata = adverse)
set.seed(30308)
crossval <- vfold_cv(training(init), strata = adverse)
```

Fit all models and estimate metrics using 10-fold cross-validation. We're not performing any tuning here (although we could do that very easily!!!).
```{r collapse = TRUE}
fit_wflows <-
  wflows %>%
  workflow_map(fn = "fit_resamples",
               seed = 30332,
               resamples = crossval,
               control = control_resamples(save_pred = TRUE),
               metrics = metric_set(roc_auc, sens, accuracy),
               verbose = TRUE)
```

Comparing our metrics for the models (unfortunately I couldn't figure out how to show which recipe was picked...)
```{r rank_baseline_models}
autoplot(fit_wflows, metric = "roc_auc")
autoplot(fit_wflows)
fit_wflows %>% collect_metrics()
```

This graph now shows which recipe was picked
```{r ae2019_improvement}
fit_wflows %>%
  collect_metrics() %>%
  separate(wflow_id, into = c("rec", "mod"), sep = "_", remove = FALSE) %>%
  ggplot(aes(x = rec, y = mean, color = mod, group = mod)) +
  geom_point() + geom_line() + facet_wrap(~ factor(.metric))
```

Looks like adding the 2019 AE didn't help much! This is evidence for our hypothesis (AE 2019 doesn't hhave much effect on the final outcome).
It's also possible that the machine learning models (svn and neural net) could benefit from some tuning.

Here are the models ranked by `roc_auc`
```{r}
fit_wflows %>% rank_results("roc_auc") %>% select(wflow_id) %>% unique()
```

We will pick `no2019ae_forest` as the "final" model.
```{r}
final_wflow <-
  fit_wflows %>%
  pull_workflow_set_result("no2019ae_forest")
```

Right now, given a test case, tries to find the probability that `ae > 3`. If that number is greater than 0.5, the model predicts `ae > 3`, if not, the model predicts `ae < 3`. This threshold of 0.5 can be changed, which will affect specificity and sensitivity. For the 10 models coming from the 10-fold CV, we compute specificity and sensitivity for all threshold in `seq(0.5, 1, 0.01)`, that is a grid from 0.5 to 1 with step of 0.01. We then average the estimate for the 10 models to get the following plot (could have also drawn error bars...)
```{r choosing_bin_threshold}
final_wflow <-
  final_wflow %>%
  rowwise() %>%
  mutate(thr_perf = list(threshold_perf(.predictions, adverse, `.pred_ae > 3`, thresholds = seq(0.5, 1, by = 0.01))))

final_wflow %>%
  select(thr_perf, id) %>%
  unnest(thr_perf) %>%
  group_by(.threshold, .metric) %>%
  summarize(estimate = mean(.estimate)) %>%
  filter(.metric != "distance") %>%
  ggplot(aes(x = .threshold, y = estimate, color = .metric)) + geom_line()
```

We can now choose the threshold based on what we need. (what do we need?)


## Tuning some models
```{r collapse = TRUE}
tune_log_spec <-
  logistic_reg(penalty = tune()) %>%
  set_engine("glmnet") %>%
  set_mode("classification")
tune_forest_spec <-
  rand_forest(trees = 1000, mtry = tune(), min_n = tune()) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
tune_sln_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine("nnet") %>%
  set_mode("classification")
tune_svm_rbf_spec <-
  svm_rbf(cost = tune(), rbf_sigma = tune(), margin = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
# tune_svm_poly_spec <-
#   svm_poly() %>%
#   set_engine("kernlab") %>%
#   set_mode("classification")
# tune_knn_spec <-
#   nearest_neighbor() %>%
#   set_engine("kknn") %>%
#   set_mode("classification")

models <- list(log = tune_log_spec,
               forest = tune_forest_spec,
               sln = tune_sln_spec,
               svm = tune_svm_rbf_spec)
recipes <- list(no2019_rec)
wflows <- workflow_set(recipes, models)

# make a bigger grid!
# or use something like finetune!
results <-
  wflows %>%
  workflow_map(resamples = crossval,
               grid = 10,
               metrics = metric_set(roc_auc, accuracy, sens, spec, ppv, npv),
               control = control_grid(save_pred = TRUE),
               seed = 828282,
               verbose = TRUE)

autoplot(results)
```

As a example, let's look at the results for the svm.
We won't touch the prediction thresholds here.
```{r}
tuned_svm <-
  results %>%
  pull_workflow_set_result("recipe_svm")


best_svm <-
  tuned_svm %>%
  select_best(metric = "roc_auc")

last_svm <-
  results %>%
  pull_workflow("recipe_svm") %>%
  finalize_workflow(best_svm) %>%
  last_fit(init,
           metrics = metric_set(roc_auc, accuracy, sens, spec, ppv, npv))

last_svm %>%
  collect_metrics()

```


