```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 6, dpi = 150, fig.path = "figures/time-"
)
```

# Time-based models
Here we will look at what happens when time-dependent data is added. At the moment, we only use deaths per day per zipcode

## Data wrangling
This is going to be mostly the same as `baseline_models.Rmd`. Refer to that file for more details.
Note that here we're not extracting the AEs.
```{r}
library(tidyverse)
library(tidymodels)
library(feather)
# library(arrow)
library(magrittr)
library(skimr)
library(lubridate)
library(timetk)
library(modeltime)
library(glue)
library(slider)
per <- read_feather("data/simulation_data/all_persons.feather")

clients <-
  per %>%
  group_by(client) %>%
  summarize(
    zip3 = first(zip3),
    size = n(),
    volume = sum(FaceAmt),
    avg_qx = mean(qx),
    avg_age = mean(Age),
    per_male = sum(Sex == "Male") / size,
    per_blue_collar = sum(collar == "blue") / size,
    expected = sum(qx * FaceAmt)
  )

zip_data <-
  read_feather("data/data.feather") %>%
  mutate(
    density = POP / AREALAND,
    AREALAND = NULL,
    AREA = NULL,
    HU = NULL,
    vaccinated = NULL,
    per_lib = NULL,
    per_green = NULL,
    per_other = NULL,
    per_rep = NULL,
    unempl_2020 = NULL,
    deaths_covid = NULL,
    deaths_all = NULL
  ) %>%
  rename(
    unemp = unempl_2019,
    hes_uns = hes_unsure,
    str_hes = strong_hes,
    income = Median_Household_Income_2019
  )

clients %<>%
  inner_join(zip_data, by = "zip3") %>%
  drop_na()
```

```{r}
deaths <-
  read_feather("data/deaths_zip3.feather") %>%
  mutate(date = ceiling_date(date, unit = "week")) %>%
  group_by(zip3, date) %>%
  summarize(totdeaths = max(deaths)) %>%
  mutate(zip_deaths = totdeaths - lag(totdeaths, default = 0)) %>%
  select(-totdeaths) %>%
  ungroup()
```

Next we look at the deaths only
```{r}
per %<>%
  drop_na() %>%
  select(-month)
```

2019 didn't have 53 weeks, so the deaths on week 53 in 2019 will be changed to death in week 1 of 2020.

Now back to the 53 week business
```{r}
wk53 <-
  per %>%
  filter(year == 2019, week == 53) %>%
  mutate(year = 2020, week = 1)
per %<>%
  rows_update(wk53, by = c("client", "participant"))
```

The tibble `per` contains all the people that die and their deathweek. Everyone should die only once now! Let's convert the week and year into the last date of that week.

```{r}
per %<>%
  nest_by(week, year) %>%
  mutate(date = ceiling_date(ymd(glue("{year}-01-01")) + weeks(week - 1), unit = "week")) %>%
  ungroup() %>%
  select(-week, -year) %>%
  unnest(cols = c(data))
```

Next, we need some kind of a rolling count for AE. Looks like the package `slider` might help.
I want actual claims per week for each client.
We note that there are 4 clients that won't have any deaths
```{r}
no_deaths <-
  clients %>%
  anti_join(per, by = "client") %>%
  select(client) %>%
  mutate(date = ceiling_date(ymd("2019-01-01"), unit = "week"), claims = 0)
```

We compute face amount per week for each client. This number is 0 if the client has no deaths that week.
```{r}
weekly_claims <-
  per %>%
  group_by(date, client) %>%
  summarize(claims = sum(FaceAmt), .groups = "drop") %>%
  bind_rows(no_deaths) %>%
  complete(date, client, fill = list(claims = 0))
```

We now have 65,369 rows, which is 131 weeks * 499 clients.

Let's merge everything.
```{r}
weekly_data <-
  clients %>%
  left_join(weekly_claims, by = c("client")) %>%
  relocate(date) %>%
  relocate(claims, .after = zip3) %>%
  left_join(deaths, by = c("date", "zip3")) %>%
  relocate(zip_deaths, .after = claims) %>%
  mutate(zip_deaths = replace_na(zip_deaths, 0), ae = claims / (expected / 52.18))
```

We add a rolling AE number. We will smooth the actual claims number of each week by taking a weighted average of actual claims in the 13 weeks prior. The weights come from a Gaussian distribution...
The function `sliding_smoother` takes a vector and outputs a vector of smoothed values.

The below picture show what the smoother does for claims of client 7. We also show a 3 month mean AE
```{r}
smoother <- function(x) { weighted.mean(x, dnorm(seq(-1, 0, length.out = length(x)), sd = 0.33)) }
sliding_smoother <-
  slidify(smoother, .period = 13, .align = "right")

sliding_mean <-
  slidify(mean, .period = 13, .align = "right")


weekly_data %>%
  filter(client == 7) %>%
  ggplot(aes(x = date)) +
  # geom_line(aes(y = sliding_mean(claims)), color = "red") +
  # geom_line(aes(y = smooth_vec(claims, period = 13)), color = "blue") +
  geom_line(aes(y = sliding_smoother(ae)), color = "magenta") +
  geom_line(aes(y = sliding_mean(ae)), color = "red") +
  geom_line(aes(y = ae), linetype = 2)
```

We add a column called `smoothed_ae` and `smoothed_deaths`.
```{r}
weekly_data <-
  weekly_data %>%
  group_by(client) %>%
  mutate(smoothed_ae = sliding_smoother(ae), smoothed_deaths = sliding_smoother(zip_deaths), .before = size) %>%
  drop_na()
```

We plot the average AE for each week. This is pretty crazy...
```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(avg_ae = mean(smoothed_ae), sd = sd(smoothed_ae)) %>%
  ggplot(aes(x = date, y = avg_ae)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed")
  # geom_errorbar(aes(ymin = avg_ae - sd, ymax = avg_ae + sd))
```

```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(avg_deaths = mean(zip_deaths / POP)) %>%
  ggplot(aes(x = date, y = avg_deaths)) + geom_line()
```

Let compute the 25th percentile AE of each week. We will pick the threshold for `adverse` based on this.
```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(
      `12.5` = quantile(smoothed_ae, 0.125),
      `25` = quantile(smoothed_ae, 0.25),
      `50` = quantile(smoothed_ae, 0.50)
  ) %>%
  pivot_longer(-date, names_to = "pth", values_to = "smoothed_ae") %>%
  ggplot(aes(x = date, y = smoothed_ae, color = pth)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed")
```


Maybe there is a better number to look at than smoothed ae. Let's shrink smoothed ae based on the log(Volume * average qx). This gives us some kind of a measure of client size and mortality.
```{r}
client_shrinkage <-
  weekly_data %>%
  summarize(dep_var = first(volume * avg_qx)) %>%
  mutate(shrinkage = rescale(log(dep_var), to = c(0.3, 1)))

ggplot(client_shrinkage, aes(x = dep_var)) + geom_density() + scale_x_log10()

processed_data <-
  weekly_data %>%
  left_join(client_shrinkage, by = "client") %>%
  ungroup() %>%
  mutate(shrunk_ae = smoothed_ae * shrinkage, .after = smoothed_ae)

processed_data %>%
  group_by(date) %>%
  summarize(avg_ae = mean(shrunk_ae)) %>%
  ggplot(aes(date, avg_ae)) + geom_line()


processed_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(
      `12.5` = quantile(shrunk_ae, 0.125),
      `25` = quantile(shrunk_ae, 0.25),
      `50` = quantile(shrunk_ae, 0.50)
  ) %>%
  pivot_longer(-date, names_to = "pth", values_to = "value") %>%
  ggplot(aes(x = date, y = value, color = pth)) +
  geom_line() +
  geom_hline(yintercept = 2.5, linetype = "dashed")
```




We will choose "smoothed, shrunk 3 month AE" > 2.5 as "Adverse".
```{r}
processed_data <-
  processed_data %>%
  ungroup() %>%
  mutate(class = factor(if_else(shrunk_ae > 2.5, "Adverse", "Not adverse"), levels = c("Adverse", "Not adverse")), .after = shrunk_ae)
```


How many clients are adverse each week?
```{r}
processed_data %>%
  group_by(date) %>%
  summarize(prop_adverse = sum(class == "Adverse") / n()) %>%
  ggplot(aes(x = date, y = prop_adverse)) + geom_line()
```

# Modeling (with tidyverts)

## Case study: Jan 1st 2021

### ARIMA on zip_deaths + random forest

We will travel back in time. We will attempt to predict six months worth of "averse" variables for all clients.
```{r}
library(tsibble)
library(fable)

data_tsibble <-
  processed_data
  # mutate(yw = yearweek(date), .before = date, .keep = "unused")

# splits <-
#   data_tsibble %>%
#   filter(date >= "2020-03-15") %>%
#   nest_by(date) %>%
#   sliding_index(index = date,
#                 lookback = weeks(12),
#                 assess_stop = weeks(26))

# split <- splits %>% pluck(1, 3)
# train <- training(split) %>% unnest(data)
# test <- testing(split) %>% unnest(data)
train_old <-
  processed_data %>%
  filter(date <= "2021-01-01")

test_old <-
  processed_data %>%
  filter(date > "2021-01-01" & date <= "2021-06-01")
```


Here we will manually forecast future `zip_deaths` using a fully default ARIMA forecaster. Maybe we should actually predict a smoothed version of `zip_deaths`???
We plot total deaths vs total forecasted deaths (in the zip codes of our clients). It's not great but it's something...
```{r cache = TRUE}
forecast <-
  train_old %>%
  filter(date >= "2020-03-15") %>%
  as_tsibble(index = date, key = client) %>%
  model(arima = ARIMA(smoothed_deaths)) %>%
  forecast(h = "4 months")
```

foo <-
  forecast %>%
  index_by(date) %>%
  summarize(fc_deaths = sum(.mean))

bar <-
  test %>%
  as_tsibble(key = client, index = date) %>%
  index_by(date) %>%
  summarize(true_deaths = sum(zip_deaths))

foo %>%
  left_join(bar, by = "date") %>%
  ggplot(aes(x = date)) + geom_line(aes(y = fc_deaths), color = "red") + geom_line(aes(y = true_deaths))


Common recipe for all models. 

```{r}
common_recipe <-
  recipe(class ~ ., data = processed_data) %>%
  step_rm(client, zip3, claims, smoothed_ae, shrunk_ae, shrinkage, dep_var, ae, smoothed_deaths, date) %>%
  step_zv(all_predictors()) %>%
  # step_log(volume, POP) %>%
  step_normalize(all_predictors())
  
```

### Multiple models

```{r}
forest_spec <-
  rand_forest(trees = 1000) %>%
  set_engine("ranger", num.threads = 8, seed = 123456789) %>%
  set_mode("classification")

log_spec <- 
  logistic_reg(
  mode = "classification",
  engine = "glm")

svm_lin_spec <-
  svm_linear() %>%
  set_engine("LiblineaR") %>%
  set_mode("classification")

knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")

sln_spec <-
  mlp(activation = "relu", hidden_units = 6, epochs = 100) %>%
  set_engine("keras", verbose=0) %>%
  set_mode("classification")


bt_spec <- boost_tree(
  mode = "classification",
  engine = "xgboost",
  trees = 100)

```

```{r}
#models <- list(log = log_spec,
 #           forest = forest_spec,
  #             neural = sln_spec,
   #            svm_linear = svm_lin_spec,
    #           boosted_trees = bt_spec,
     #          knn = knn_spec)
#recipes <- list(common_recipe)
#wflows <- workflow_set(recipes, models)
```

```{r}
#wflows %>% 
 # mutate(fit_wflow = map(wflow_id, ~pull_workflow(.,.x)))

```

```{r}
#trained_wflows_fit <- 
  
  #wflows %>% 
 # fit(train)
```

```{r}
bt_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(bt_spec)

log_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(log_spec)

forest_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(forest_spec)

svm_lin_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(svm_lin_spec)

knn_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(knn_spec)

sln_wf <-
  workflow() %>%
  add_recipe(common_recipe) %>%
  add_model(sln_spec)
```


```{r}
wflows <- tribble(~wflow ,
                 # sln_wf,
                  knn_wf, log_wf, forest_wf, bt_wf,# svm_lin_wf
)
 


wflows <-
  wflows %>%
  mutate(wflows_fit = map(wflow, ~ fit(.x, train_old)))
```

```{r}
wflows <-
  wflows %>%
  mutate(
    class_predict = map(wflows_fit, ~ predict(.x, test_old),
    prob_predict = map(wflows_fit, ~ predict(.x, test_old, type = "prob"))))
```    

```{r}
wflows %>%
  bind_cols(tribble(~id, "knn", "log", "forest", "bt")) %>%
  select(-wflow, -wflows_fit) %>%
  mutate(prob_predict = map(prob_predict, ~ bind_cols(.x, test %>% select(date, class)))) %>%
  unnest(c(class_predict, prob_predict)) %>%
  group_by(id, date) %>%
  summarize(
            sens = sens_vec(class, .pred_class),
            spec = spec_vec(class, .pred_class),
            roc_auc = roc_auc_vec(class, .pred_Adverse), .groups = "keep") %>%
  pivot_longer(sens:roc_auc, names_to = "metric", values_to = "value") %>%
  ungroup() %>%
  ggplot(aes(x = date, y = value, color = id)) +
  geom_point() +
  geom_line() +
  facet_wrap( ~ metric)

```


First, we forecast the deaths. 
```{r}
forecasted_test <-
  forecast %>%
  as_tibble() %>%
  select(client, date, .mean) %>%
  right_join(test, by = c("client", "date")) %>%
  select(-zip_deaths) %>%
  rename(zip_deaths = .mean)

```

