# Baseline model

The goal of this documents is to compare some tuned and untuned models on a "baseline" dataset. We want to predict whether AE in 2020 goes above 3. We are only using 2019 data here! Next steps will be adding time-dependent data (e.g. COVID deaths per week per zipcode).

## Common data
```{r}
library(tidyverse)
library(tidymodels)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
per <- read_feather("data/simulation_data/all_persons.feather")
```

Compute some summary statistic for each client.
```{r}
clients <-
  per %>%
  group_by(client) %>%
  summarize(
    zip3 = first(zip3),
    size = n(),
    volume = sum(FaceAmt),
    avg_qx = mean(qx),
    avg_age = mean(Age),
    per_male = sum(Sex == "Male") / size,
    per_blue_collar = sum(collar == "blue") / size,
    expected = sum(qx * FaceAmt),
    actual_2020 = sum(FaceAmt[year == 2020], na.rm = TRUE),
    ae_2020 = actual_2020 / expected,
    actual_2019 = sum(FaceAmt[year == 2019], na.rm = TRUE),
    ae_2019 = actual_2019 / expected,
    adverse = as_factor(if_else(ae_2020 > 3, "ae > 3", "ae < 3"))
  ) %>%
  relocate(adverse, ae_2020, .after = zip3) %>%
  mutate(adverse = fct_relevel(adverse, c("ae > 3", "ae < 3")))
```

We can add some demographic information based on zip3.
```{r}
zip_data <-
  read_feather("data/data.feather") %>%
  mutate(
    density = POP / AREALAND,
    AREALAND = NULL,
    AREA = NULL,
    HU = NULL,
    vaccinated = NULL,
    per_lib = NULL,
    per_green = NULL,
    per_other = NULL,
    per_rep = NULL,
    unempl_2020 = NULL,
    deaths_covid = NULL,
    deaths_all = NULL
  ) %>%
  rename(
    unemp = unempl_2019,
    hes_uns = hes_unsure,
    str_hes = strong_hes,
    income = Median_Household_Income_2019
  )
```
There seems to be some clients with some zip codes that we cannot deal with. These are the ones
```{r}
clients %>%
  anti_join(zip_data, by = "zip3") %>%
  select(zip3)
```
These correspond to the following areas

ZIP3 | Area       |
-----|------------|
969  | Guam, Palau, Federated States of Micronesia, Northern Mariana Islands, Marshall Islands |
093  | Military bases in Iraq and Afghanistan |
732  | Not in use |
872  | Not in use |
004  | Not in use |
202  | Washington DC, Government 1 |

We ignore clients with these zip codes. There are also two clients in DC for which we're missing election data. We will ignore those as well.
```{r}
clients %<>%
  inner_join(zip_data, by = "zip3") %>%
  drop_na()
```

We now have our full dataset. Behold!
```{r}
skim(clients)
```

## Workflow set
We'll evaluate models using a workflow set. To make our life easier, we will remove some variables and use a formula instead of a recipe.
```{r}
clients <-
  clients %>%
  select(-client, -zip3, -ae_2020, -actual_2020, -actual_2019)
```

We now gather our recipes and models.
```{r}
with2019_rec <-
  recipe(adverse ~ ., data = clients) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors(), -all_nominal())
no2019_rec <-
  with2019_rec %>%
  step_rm(ae_2019)

log_spec <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
tuned_log_spec <-
  logistic_reg(penalty = 0.00118) %>%
  set_engine("glmnet") %>%
  set_mode("classification")
forest_spec <-
  rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
tuned_forest_spec <-
  rand_forest(trees = 1000, mtry = 12, min_n = 21) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
sln_spec <-
  mlp() %>%
  set_engine("nnet") %>%
  set_mode("classification")
svm_rbf_spec <-
  svm_rbf() %>%
  set_engine("kernlab") %>%
  set_mode("classification")
knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")

# Tuned models
tuned_sln_spec <-
  mlp(penalty=1.023531e-09) %>%
  set_engine("nnet") %>%
  set_mode("classification")
tuned_svm_rbf_spec <-
  svm_rbf(cost=4.350593e+00) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
tuned_knn_spec <-
  nearest_neighbor(neighbors=6) %>%
  set_engine("kknn") %>%
  set_mode("classification")


models <- list(log = log_spec,
               logtuned = tuned_log_spec,
               forest = forest_spec,
               foresttuned = tuned_forest_spec,
               neural = sln_spec,
               neuraltuned=tuned_sln_spec,
               svmrbf = svm_rbf_spec,
               svmrbftuned=tuned_svm_rbf_spec,
               knnspec = knn_spec,
               knnspectuned= tuned_knn_spec)
recipes <- list("with2019ae" = with2019_rec,
                "no2019ae" = no2019_rec)
wflows <- workflow_set(recipes, models)
```

Data splitting
```{r}
set.seed(30308)
init <- initial_split(clients, strata = adverse)
set.seed(30308)
crossval <- vfold_cv(training(init), strata = adverse)
```

Fit all models and estimate metrics using 10-fold cross-validation. We're not performing any tuning here (although we could do that very easily!!!).
```{r collapse = TRUE}
fit_wflows <-
  wflows %>%
  workflow_map(fn = "fit_resamples",
               seed = 30332,
               resamples = crossval,
               control = control_resamples(save_pred = TRUE),
               metrics = metric_set(roc_auc, sens, accuracy, spec, npv, ppv),
               verbose = TRUE)
```

Comparing our metrics for the models (unfortunately I couldn't figure out how to show which recipe was picked...)
```{r rank_baseline_models}
autoplot(fit_wflows, metric = "roc_auc")
autoplot(fit_wflows)
fit_wflows %>% collect_metrics()
```

This graph now shows which recipe was picked
```{r ae2019_improvement}
fit_wflows %>%
  collect_metrics() %>%
  separate(wflow_id, into = c("rec", "mod"), sep = "_", remove = FALSE) %>%
  ggplot(aes(x = rec, y = mean, color = str_replace(mod,"tuned",""), linetype=str_detect(mod,"tuned"), group = mod)) +
  geom_point() + geom_line() + facet_wrap(~ factor(.metric))
```



Here are the models ranked by `roc_auc`
```{r}
fit_wflows %>% rank_results("roc_auc") %>% select(wflow_id) %>% unique()
```
```{r}
fit_wflows %>% rank_results("sens") %>% select(wflow_id) %>% unique()
```


