```{r, setup, include=FALSE}
knitr::opts_chunk$set(
      fig.width = 5, fig.height = 5, dpi = 300, fig.path = "figures/final-"
)
```

# FINAL PRESENTATION

## Data loading and boring stuff
No need to talk about this.
Please download the latest version of `data/processed_data_20_12_23.feather` from Google Drive.
```{r message = FALSE}
library(tidyverse)
library(tidymodels)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
library(ggbeeswarm)
library(finetune)
library(lubridate)
library(glue)
library(slider)
processed_data <-
  read_feather("data/processed_data_20_12_23.feather")
```





# Introduction
AEs changing in 2019, 2020, copy paste from pres2

We explain how we choose the threshold for adverse and not
```{r}
processed_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(
      `12.5` = quantile(shrunk_ae, 0.125),
      `25` = quantile(shrunk_ae, 0.25),
      `50` = quantile(shrunk_ae, 0.50)
  ) %>%
  pivot_longer(-date, names_to = "pth", values_to = "value") %>%
  ggplot(aes(x = date, y = value, color = pth)) +
  geom_line() +
  geom_hline(yintercept = 2.5, linetype = "dashed")
```

How many clients are adverse each week?
```{r}
processed_data %>%
  group_by(date) %>%
  summarize(prop_adverse = sum(class == "Adverse") / n()) %>%
  ggplot(aes(x = date, y = prop_adverse)) + geom_line()
```
# "Baseline" stuff
Can maybe also copy paste some things from pres 2.

This is trained on known COVID-19 data.

* Why? It can be used as is in a future pandemic.
* Why not? Future pandemics may have different target mortality compared to COVID19.

# Time stuff

* Why? This model is more reactive than the previous one. It's trained on current pandemic data. It can be updated with time.
* Why not? This will be very bad early on in the pandemic (lack of data) and bad when forecasting is hard.

## Model selection
Looking at performance every day for 3 months in the future

## Training on part of the data (forecasted smoothed deaths + IHME)
Testing on known clients and unknown clients.

### Case study: June 2020

### Case study: Jan 2021

## Predicting AE as a time-series
Don't take weekly deaths into account. How well/badly can we do?
