```{r, setup, include=FALSE}
knitr::opts_chunk$set(
      fig.width = 5, fig.height = 5, dpi = 300, fig.path = "figures/pres-"
)
```


# Presentation 2 script

## Data loading
No need to talk about this.
Please download the latest version of `data/simulation_data/all_persons.feather` from Google Drive.
```{r message = FALSE}
library(tidyverse)
library(tidymodels)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
library(ggbeeswarm)
library(finetune)
per <- read_feather("data/simulation_data/all_persons.feather")

clients <-
  per %>%
  group_by(client) %>%
  summarize(
    zip3 = first(zip3),
    size = n(),
    volume = sum(FaceAmt),
    avg_qx = mean(qx),
    avg_age = mean(Age),
    per_male = sum(Sex == "Male") / size,
    per_blue_collar = sum(collar == "blue") / size,
    expected = sum(qx * FaceAmt),
    actual_2021 = sum(FaceAmt[year == 2021], na.rm = TRUE),
    ae_2021 = actual_2021 / (expected / 2),
    actual_2020 = sum(FaceAmt[year == 2020], na.rm = TRUE),
    ae_2020 = actual_2020 / expected,
    actual_2019 = sum(FaceAmt[year == 2019], na.rm = TRUE),
    ae_2019 = actual_2019 / expected,
    adverse = as_factor(if_else(ae_2020 > 3, "ae > 3", "ae < 3"))
  ) %>%
  relocate(adverse, ae_2020, .after = zip3) %>%
  mutate(adverse = fct_relevel(adverse, c("ae > 3", "ae < 3")))


zip_data <-
  read_feather("data/data.feather") %>%
  mutate(
    density = POP / AREALAND,
    AREALAND = NULL,
    AREA = NULL,
    HU = NULL,
    vaccinated = NULL,
    per_lib = NULL,
    per_green = NULL,
    per_other = NULL,
    per_rep = NULL,
    unempl_2020 = NULL,
    deaths_covid = NULL,
    deaths_all = NULL
  ) %>%
  rename(
    unemp = unempl_2019,
    hes_uns = hes_unsure,
    str_hes = strong_hes,
    income = Median_Household_Income_2019
  )


clients %<>%
  inner_join(zip_data, by = "zip3") %>%
  drop_na()
```

We now have our full dataset. Behold!
```{r}
skim(clients)
```

## How many were actually adverse?
We count how many have AE > 1 in each year
```{r}
clients %>%
  transmute(
    `2019` = ae_2019 > 1,
    `2020` = ae_2020 > 1,
    `2021` = ae_2021 > 1) %>%
  pivot_longer(`2019`:`2021`, names_to = "year", values_to = "adverse") %>%
  mutate(adverse = fct_rev(fct_recode(factor(adverse), `AE > 1` = "TRUE", `AE < 1` = "FALSE"))) %>%
  ggplot(aes(x = year, fill = adverse)) + geom_bar() +
  labs( x = "Year", y = "Count", fill = "Class", title = "Number of clients experiencing adverse deaths")
  # scale_fill_manual(values = c("yellow2", "deepskyblue"))
```

Changes in actual claims from 2019 to 2021. Here each point is a company. Note the y-axis is logarithmic!!!
```{r}
set.seed(92929292)
clients %>%
  select(expected, actual_2019, actual_2020, actual_2021) %>%
  rename(actual_Expected = expected) %>%
  pivot_longer(everything(), names_to = "Year", values_to = "Claims") %>%
  mutate(Year = str_sub(Year, 8)) %>%
  filter(Claims > 0) %>%
  ggplot(aes(Year, Claims, color = Year)) + scale_y_log10() + geom_beeswarm(size = 0.5, priority = "random") +
  guides(color = "none") + labs(title = "Size of claims") +
  scale_color_manual(values = c("yellow3", "deepskyblue", "black", "red"))
```

```{r}
ggplot(clients) +
  geom_density(aes(x = actual_2019), fill = "deepskyblue", alpha = 0.5) +
  geom_density(aes(x = expected), fill = "black", alpha = 0.5) +
  scale_x_log10()
  #geom_point(data=data, aes(x=client, y=expected), colour="deepskyblue")+
  #geom_point(data=data, aes(x=client, y=actual2021), color="black") +
 # scale_y_continuous(name="Actual vs Expected Claims in 2019")
```
```{r}
ggplot(clients) +
  geom_density(aes(x = actual_2020), fill = "deepskyblue", alpha = 0.5) +
  geom_density(aes(x = expected), fill = "black", alpha = 0.5) +
  scale_x_log10()
```
```{r}
ggplot(clients) +
  geom_density(aes(x = actual_2021), fill = "deepskyblue", alpha = 0.5) +
  geom_density(aes(x = expected), fill = "black", alpha = 0.5) +
  scale_x_log10()
```
##Trying many models with and without 2019 AE as a predictor

```{r}
clients <-
  clients %>%
  select(
         -ae_2020, -ae_2021, -actual_2020, -actual_2019, -actual_2021,
         -hes, -hes_uns, -str_hes)


with2019_rec <-
  recipe(adverse ~ ., data = clients) %>%
  update_role(zip3, client, new_role = "ID") %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors(), -all_nominal())
no2019_rec <-
  with2019_rec %>%
  step_rm(ae_2019)

log_spec <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
tuned_log_spec <-
  logistic_reg(penalty = 0.00118) %>%
  set_engine("glmnet") %>%
  set_mode("classification")
forest_spec <-
  rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
tuned_forest_spec <-
  rand_forest(trees = 1000, mtry = 12, min_n = 21) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
sln_spec <-
  mlp() %>%
  set_engine("nnet") %>%
  set_mode("classification")
svm_rbf_spec <-
  svm_rbf() %>%
  set_engine("kernlab") %>%
  set_mode("classification")
svm_poly_spec <-
  svm_poly() %>%
  set_engine("kernlab") %>%
  set_mode("classification")
knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")

models <- list(log = log_spec,
               logtuned = tuned_log_spec,
               forest = forest_spec,
               foresttuned = tuned_forest_spec,
               neural = sln_spec,
               svmrbf = svm_rbf_spec,
               svmpoly = svm_poly_spec,
               knnspec = knn_spec)
recipes <- list("with2019ae" = with2019_rec,
                "no2019ae" = no2019_rec)
wflows <- workflow_set(recipes, models)
```
```{r}
set.seed(30308)
init <- initial_split(clients, strata = adverse)
set.seed(30308)
crossval <- vfold_cv(training(init), strata = adverse)
```

```{r mod_with_without_2019, collapse = TRUE, cache = TRUE}
fit_wflows <-
  wflows %>%
  workflow_map(fn = "fit_resamples",
               seed = 30332,
               resamples = crossval,
               control = control_resamples(save_pred = TRUE),
               metrics = metric_set(roc_auc, sens, accuracy),
               verbose = TRUE)

fit_wflows %>%
  collect_metrics() %>%
  filter(.metric %in% c("roc_auc", "accuracy")) %>%
  separate(wflow_id, into = c("rec", "mod"), sep = "_", remove = FALSE) %>%
  ggplot(aes(x = rec, y = mean, color = mod, group = mod)) +
  geom_point() + geom_line() + facet_wrap(~ factor(.metric)) +
  labs(color = "Model", x = NULL, y = "Value", title = "Performance of models with/without 2019 data")
```





## Trying many models with many tuning parameters

```{r}
tune_log_spec <-
  logistic_reg(penalty = tune()) %>%
  set_engine("glmnet") %>%
  set_mode("classification")
tune_forest_spec <-
  rand_forest(trees = 1000, mtry = tune(), min_n = tune()) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
tune_sln_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine("nnet") %>%
  set_mode("classification")
tune_svm_rbf_spec <-
  svm_rbf(cost = tune(), rbf_sigma = tune(), margin = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
# tune_svm_poly_spec <-
#   svm_poly() %>%
#   set_engine("kernlab") %>%
#   set_mode("classification")
tune_knn_spec <-
  nearest_neighbor(neighbors = tune(), dist_power = tune()) %>%
  set_engine("kknn") %>%
  set_mode("classification")

models <- list(log = tune_log_spec,
               forest = tune_forest_spec,
               sln = tune_sln_spec,
               svm = tune_svm_rbf_spec,
               knn = tune_knn_spec)
recipes <- list(no2019_rec)
wflows <- workflow_set(recipes, models)
```

```{r tune_many_models, cache = TRUE, collapse = TRUE}
# make a bigger grid!
# or use something like finetune!
results <-
  wflows %>%
  workflow_map(resamples = crossval,
               grid = 10,
               metrics = metric_set(roc_auc, accuracy),
               control = control_grid(save_pred = TRUE),
               seed = 828282,
               verbose = TRUE)
```
```{r fig.width = 8, fig.height = 4, dpi = 450}
autoplot(results)
```

## Tuning a forest

We don't need to normalize the predictors
```{r collapse = TRUE}
forest_spec <-
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 654)

forest_rec <-
  recipe(adverse ~ ., data = clients) %>%
  step_rm(ae_2019) %>%
  update_role(client, zip3, new_role = "ID") %>%
  step_zv(all_predictors())

forest_wflow <-
  workflow() %>%
  add_model(forest_spec) %>%
  add_recipe(no2019_rec)

forest_params <-
  forest_wflow %>%
  parameters() %>%
  update(mtry = mtry(c(1, 20)))

forest_grid <-
  grid_regular(forest_params, levels = 10)
```
```{r tune_forest, cache = TRUE, collapse = TRUE}
forest_tune <-
  forest_wflow %>%
  tune_grid(
      resamples = crossval,
      grid = forest_grid,
      metrics = metric_set(roc_auc, accuracy),
      control = control_grid(verbose = FALSE)
  )
```
```{r dpi = 450}
forest_tune %>%
  show_best(metric = "roc_auc", n = 10)

autoplot(forest_tune)
```

## These are our optimal parameters


```{r}
best_params <- list(mtry = 5, min_n = 6)
```

## Threshold the forest

use library `probably`

```{r fig.width = 7}
forest_resamples <-
  forest_wflow %>%
  finalize_workflow(best_params) %>%
  fit_resamples(
      resamples = crossval,
      control = control_resamples(save_pred = TRUE)
  )

forest_resamples %>%
  select(.predictions) %>%
  unnest(.predictions) %>%
  roc_curve(adverse, `.pred_ae > 3`) %>%
  autoplot()

forest_resamples <-
  forest_resamples %>%
  rowwise() %>%
  mutate(thr_perf = list(threshold_perf(.predictions, adverse, `.pred_ae > 3`, thresholds = seq(0.0, 1, by = 0.01))))



forest_resamples %>%
  select(thr_perf, id) %>%
  unnest(thr_perf) %>%
  group_by(.threshold, .metric) %>%
  summarize(estimate = mean(.estimate)) %>%
  filter(.metric != "distance") %>%
  ggplot(aes(x = .threshold, y = estimate, color = .metric)) + geom_line() +
  geom_vline(xintercept = 0.67, linetype = "dashed") +
  labs(x = "Estimate", y = "Threshold", color = "Metric", title = "Sensitivity and specificity by threshold")
```

We select 0.67 as our final threshold
```{r}
my_threshold <- 0.67
```


## This is our final model !!!
```{r}
final_fit <-
  forest_wflow %>%
  finalize_workflow(best_params) %>%
  last_fit(init)

trained_wflow <-
  final_fit %>%
  extract_workflow()

trained_recipe <-
  trained_wflow %>%
  extract_recipe(estimated = TRUE)
```


## Apply chosen threshold and compute metrics
Use `probably` to apply threshold.
Conf. matrix, accuracy, sens, spec, whatever you like

```{r}
final_fit %>%
  collect_metrics()

thresholded_predictions <-
  final_fit %>%
  collect_predictions() %>%
  select(-.pred_class) %>%
  mutate(class_pred = 
            make_two_class_pred(
                  `.pred_ae > 3`,
                  levels = levels(clients$adverse),
                  threshold = my_threshold))

confusion_matrix <-
  thresholded_predictions %>%
  conf_mat(adverse, class_pred)

confusion_matrix %>%
  autoplot(type = "heatmap")

confusion_matrix %>% summary()
```

#SHAP value is not done yet. It is done if we can take the whole clients set (not dividing between train and test). But if we want to use it on test only, needs more work. (see pictures on slack for the whole client set). 

```{r}
fulltest <- testing(init)
fulltrain <- training(init)
```

```{r eval = FALSE}
library(DALEX)
fit_parsnip <- trained_wflow %>% extract_fit_parsnip
train <- trained_recipe %>% bake(training(init))
test <- trained_recipe %>% bake(testing(init))
ex <-
  explain(
    model = fit_parsnip,
    data = train)

ex %>%
  predict_parts(train %>% slice(343)) %>%
  plot()

fit_parsnip %>% predict(train %>% slice(343), type = "prob")
```

```{r}
shap <- predict_parts(explainer = ex, 
                      new_observation = train%>%slice(343), 
                                 type = "shap",
                                  B = 25)
                               
```

```{r}
plot(shap,show_boxplots = FALSE)
```

```{r eval = FALSE}
library(treeshap)
```

```{r eval = FALSE}
clients$adverse <- ifelse(clients$adverse == "ae > 3", 1, 0)
```

```{r eval = FALSE}
model <- final_fit %>%
  extract_fit_engine()
```



```{r eval = FALSE}
test <- testing(init)
test$adverse <- ifelse(test$adverse == "ae > 3", 1, 0)
test
```

```{r eval = FALSE}
final_pred
test
test_pred <- merge((final_pred, test), 
```

```{r eval = FALSE}
model_unified <- ranger.unify(model, )
```


#autoplot(cm, type = "heatmap")


