```{r, setup, include=FALSE}
knitr::opts_chunk$set(
      fig.path = "figures/report/fig-"
)
```


# Executive summary

# Data wrangling

In this section, we describe our data gathering and tidying process.
We will be making extensive use of the `tidyverse` family of packages.
A series of scripts are used to generate tibbles, which are then saved in a `*.feather` for fast loading.
A full list of scripts can be viewed in the Appendix (TODO!!!)
```{r message = FALSE}
library(tidyverse)
library(feather)
```

## Data sources

Our dataset consists of two parts: publicly obtained data and simulated clients.

filename | source | description
---------|--------|------------
`covid_deaths_usafacts.csv` | [USAFacts](https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/) | Cumulative weekly COVID-19 deaths by county
`soa_base_2017.csv` | (Sent by Douglas Armstrong) | $q_x$ values by gender, age, industry


### US Census bureau
We used the US Census Bureau's API to obtain the 2019 estimates for population and density per county from the Census Bureau's Population Estimates Program (PEP).
The `censusapi` package provides an R interface to the API.
Using the API requires an API key, which can be obtained from [here](https://api.census.gov/data/key_signup.html).
The following snippet fetches the data, and saves the tibble into a file called `pop_den.feather`. See also `data/census.R`.
```{r eval = FALSE}
library("censusapi")

Sys.setenv(CENSUS_KEY = "YOUR_KEY_HERE")

# date_code = 12 is an estimate for July 1, 2019
# total population + density
pop <- getCensus(
                 name = "pep/population",
                 vintage = 2019,
                 region = "county:*",
                 vars = c("POP", "DENSITY"),
                 DATE_CODE = 12)
pop <- tibble(pop) %>%
  select(-DATE_CODE)
write_feather(pop, "pop_den.feather")
```

See also `data/all_persons.r`.


## County to zip3

## Weekly deaths
See also `data/deaths.R`.

## IHME (state to zip3)

## Simulated client dataset
The clients we were tasked to study were simulated by Securian Financial.
The dataset consists of 20 files called `data/simulation_data/experience_weekly_{n}.RDS` and `data/simulation_data/person_{n}.RDS` for $n = 1,\dotsc, 10$.
In total, we have 500 clients and 1,382,321 individuals.

The `person_{n}.RDS` files contain information such as company, zip code, age, face amount, gender, and collar (blue or white, but in this dataset every indivual was blue collar).
The rows in `experience_weekly_{n}.RDS` correspond to individuals and weeks, and contains in a flag `deaths` that becomes 1 on the week they die.
In total, these tables contain 170,025,483 rows, but the same information can be conveyed in 1,382,231 rows by attaching to each individual their death date (or `NA` if they don't die).
```{r eval = FALSE}
read_data <- function(n) {
  exp_name <- str_glue("simulation_data/experience_weekly_{n}.RDS")
  per_name <- str_glue("simulation_data/person_{n}.RDS")
  exp <- read_rds(exp_name)
  per <- read_rds(per_name)

  dies <-
    exp %>%
    filter(death > 0) %>%
    select(client, participant, week, month, year)
  aug_per <-
    per %>%
    left_join(dies, by = c("client", "participant"))

  aug_per
}

all_persons <- (1:10) %>% map_dfr(read_data)
```

We noticed that some individuals die more than once. This removes multiple deaths
```{r eval = FALSE}
all_persons <-
  all_persons %>%
  group_by(client, participant) %>%
  arrange(year, week, .by_group = TRUE) %>%
  slice_head()
```

We finally attach to each individual their yearly $q_x$ value, and save the resuilting tibble in `data/simultation_data/all_persons.feather`.

```{r eval = FALSE}
qx_table <- read_csv("soa_base_2017.csv")

all_persons <-
  all_persons %>%
  left_join(qx_table, by = c("Age", "Sex", "collar")) %>%
  relocate(qx, .after = collar)

write_feather(all_persons %>% ungroup(), "simulation_data/all_persons.feather")
```

## Other cleanup
Bad zipcodes

## Final merged tibble

# Long-term model

## Intro

## Methods

## Results


# Short-term model

## Intro

## Methods

## Results

# Conclusion

# Appendices

## Data repository

## R scripts

## Rmd files
