
## Common data
```{r}
library(tidyverse)
library(tidymodels)
library(modeltime)
library(timetk)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
library(dplyr)
library(lubridate)
```



Get weeklydata from time.Rmd
```{r}
clients<-read_feather("process.feather")%>%
  filter(date >= "2020-03-15")%>%
  mutate(client = as.factor(client))%>%
  #select(-year)%>%
  mutate(POP=log(POP))%>%
  mutate(volume = log(volume))%>%
  mutate(expected = log(expected))
```

split
```{r}
#split
  split
  set.seed = 1234
  splits <-  clients %>% time_series_split(initial = "6 months", assess = "6 months", date_var = date, cumulative = TRUE)
  train = training(splits)
  test = testing(splits)
  #index <- unique(train$client)
  #val <- sliding_index(training(splits), index)
  
  #crossval <- vfold_cv(training(splits), strata = zip3)
```
#window forest
It is not good 
```{r}
model_spec <- window_reg(
        window_size     = "6 months",
        id = "client"
    ) %>%
    # Extra parameters passed as: set_engine(...)
    set_engine(
        engine          = "window_function",
        #window_function = median,
        na.rm           = TRUE,
        window_function = ~ tail(.x, 100),
    )

# Fit Spec
model_fit <- model_spec %>%
    fit(shrunk_ae~ date + client, data = train)
model_fit
predict(model_fit, test)


model_tbl <- modeltime_table(
    model_fit)

model_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) %>%
    group_by(client) %>%
    filter(client == c( "1", "5", "7"))%>%
    plot_modeltime_forecast(
        .facet_ncol  = 3,
        .interactive = FALSE,
        .title = "Forecast Plot",
        .line_alpha = 0.6,
        .line_size = 1,
        .y_intercept = 3.0
    )
calib_tbl <- model_tbl %>%
    modeltime_calibrate(
      new_data = testing(splits), 
      id       = "client"
    )
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = FALSE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = TRUE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)

```

```{r}
result0 <- calib_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) 
result0
```

#Compare only one predictor(date) with all predictors(as before), the latter has better results.
#XGboost have the best result among all 7 model.

We now gather our recipes and models.

#recipe
```{r}
rec_obj <-
    recipe(shrunk_ae ~ ., data = training(splits)) %>%
    #step_rm(year,month,day)%>%
    step_rm(zip3)%>%
    step_rm( claims, smoothed_ae, class, shrinkage, dep_var, ae, zip_deaths)%>%
    #step_rm(adverse)%>%
    step_mutate(client = droplevels(client)) %>%
    step_timeseries_signature(date) %>%
    step_rm(date)%>%
    step_dummy(all_nominal_predictors(), one_hot = TRUE)%>%
     step_zv(all_predictors()) %>%
    step_normalize(all_predictors(), -all_nominal())

summary(prep(rec_obj))
bake(prep(rec_obj),training(splits))
```

#engine
```{r}
forest_spec <-
  rand_forest(trees = 1000) %>%
  set_engine("ranger", num.threads = 8, seed = 123456789) %>%
  #set_mode("classification") %>%
  set_mode("regression")%>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
tuned_forest_spec <-
  rand_forest(trees = 1000, mtry = 12, min_n = 21) %>%
  #set_mode("classification") %>%
  set_mode("regression")%>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
sln_spec <-
  mlp(activation = "relu", hidden_units = 6, epochs = 100) %>%
  set_engine("nnet") %>%
  #set_mode("classification")
  set_mode("regression")
svm_rbf_spec <-
  svm_rbf() %>%
  set_engine("kernlab") %>%
  #set_mode("classification")
  set_mode("regression")
svm_poly_spec <-
  svm_poly() %>%
  set_engine("kernlab") %>%
  #set_mode("classification")
  set_mode("regression")
knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  #set_mode("classification")
  set_mode("regression")
xgboost_spec <-
  boost_tree(trees = 100) %>%
  set_engine("xgboost") %>%
  #set_mode("classification")
  set_mode("regression")
```

# workflow without death data
```{r eval= FALSE}
#cannot fit
wflw_neural <- workflow() %>%
    add_model(
        sln_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))
```
```{r eval= FALSE}
#take a lot time()
wflw_svmpoly <- workflow() %>%
    add_model(
        svm_poly_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))
```

```{r}
wflw_rf <- workflow() %>%
    add_model(
        forest_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_tunedrf <- workflow() %>%
    add_model(
        tuned_forest_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_svmrbf <- workflow() %>%
    add_model(
        svm_rbf_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_knnspec <- workflow() %>%
    add_model(
        knn_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))  

wflw_xgboost <- workflow() %>%
    add_model(
        xgboost_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))  
```
Create a Modeltime Table
```{r}
model_tbl <- modeltime_table(
    wflw_rf,
    wflw_tunedrf,
    #wflw_neural,
    wflw_svmrbf,
    #wflw_svmpoly,
    wflw_knnspec,
    wflw_xgboost
)
model_tbl
```

#Calibrate by client
```{r}
calib_tbl <- model_tbl %>%
    modeltime_calibrate(
      new_data = testing(splits), 
      id       = "client"
    )

calib_tbl
```

Measure Accuracy on validation data
```{r}
#global error
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = FALSE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
```{r}
#local error for each client
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = TRUE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
validate the  Test Data
conf_by_id : produce confidence interval estimates by an ID feature.
#all results
```{r}
result <- calib_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) 
result
```
```{r}
result %>%
    group_by(client) %>%
    filter(client == c( "1", "5", "7","10", "61","100"))%>%
    plot_modeltime_forecast(
        .facet_ncol  = 3,
        .interactive = FALSE,
        .title = "Forecast Plot",
        .line_alpha = 0.6,
        .line_size = 1,
        .y_intercept = 3.0
    )
```

#recipe with zip death
```{r}
rec_obj1 <-
    recipe(shrunk_ae ~ ., data = training(splits)) %>%
    #step_rm(year,month,day)%>%
    step_rm(zip3)%>%
    step_rm( claims, smoothed_ae, class, shrinkage, dep_var, ae)%>%
    #step_rm(adverse)%>%
    step_mutate(client = droplevels(client)) %>%
    step_timeseries_signature(date) %>%
    step_rm(date)%>%
    step_dummy(all_nominal_predictors(), one_hot = TRUE)%>%
     step_zv(all_predictors()) %>%
    step_normalize(all_predictors(), -all_nominal())

summary(prep(rec_obj1))
bake(prep(rec_obj1),training(splits))
```


# workflow with zip death

```{r}
wflw_rf1 <- workflow() %>%
    add_model(
        forest_spec
    ) %>%
    add_recipe(rec_obj1) %>%
    fit(data = training(splits))

wflw_tunedrf1 <- workflow() %>%
    add_model(
        tuned_forest_spec
    ) %>%
    add_recipe(rec_obj1) %>%
    fit(data = training(splits))

wflw_svmrbf1 <- workflow() %>%
    add_model(
        svm_rbf_spec
    ) %>%
    add_recipe(rec_obj1) %>%
    fit(data = training(splits))

wflw_knnspec1 <- workflow() %>%
    add_model(
        knn_spec
    ) %>%
    add_recipe(rec_obj1) %>%
    fit(data = training(splits))  

wflw_xgboost1 <- workflow() %>%
    add_model(
        xgboost_spec
    ) %>%
    add_recipe(rec_obj1) %>%
    fit(data = training(splits))  
```
Create a Modeltime Table
```{r}
model_tbl1 <- modeltime_table(
    wflw_rf1,
    wflw_tunedrf1,
    #wflw_neural,
    wflw_svmrbf1,
    #wflw_svmpoly,
    wflw_knnspec1,
    wflw_xgboost1
)
model_tbl1
```

#Calibrate by client
```{r}
calib_tbl1 <- model_tbl1 %>%
    modeltime_calibrate(
      new_data = testing(splits), 
      id       = "client"
    )

calib_tbl1
```

Measure Accuracy on validation data
```{r}
#global error
calib_tbl1 %>% 
    modeltime_accuracy(acc_by_id = FALSE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
```{r}
#local error for each client
calib_tbl1 %>% 
    modeltime_accuracy(acc_by_id = TRUE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
validate the  Test Data
conf_by_id : produce confidence interval estimates by an ID feature.
#all results
```{r}
result1 <- calib_tbl1 %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) 
result1
```
```{r}
result1 %>%
    group_by(client) %>%
    filter(client == c( "1", "5", "7","10", "61","100"))%>%
    plot_modeltime_forecast(
        .facet_ncol  = 3,
        .interactive = FALSE,
        .title = "Forecast Plot",
        .line_alpha = 0.6,
        .line_size = 1,
        .y_intercept = 3.0
    )
```


#plot sens, spec, accuracy, j_index
```{r}
actual <- clients %>%
  select(date, client, shrunk_ae)
```

```{r}
predresult0 <- result0 %>%
  select(-.model_desc, -.conf_lo, -.conf_hi, -.key) %>%
  rename(model = .model_id, value = .value, date= .index)%>%
  relocate(model, value, .after = client)%>%
  pivot_wider(names_from = model, values_from =value)%>%
  rename(actual = "NA", windows = "1" )%>%
  drop_na()%>%
  mutate(obs =  ifelse(actual > 2.5, TRUE, FALSE))%>%
  mutate(windows_pred=  ifelse(windows > 2.5,TRUE, FALSE))%>%
  mutate(obs = as.factor(obs), windows_pred = as.factor(windows_pred))

conf0<-predresult0%>%
  group_by(date) %>%
  summarize(sens_windows = sens_vec(obs, windows_pred),
            spec_windows = spec_vec(obs, windows_pred), 
            jinex_window = j_index_vec(obs,windows_pred),
            acc_windows = accuracy_vec(obs, windows_pred))

conf0%>%
  pivot_longer(sens_windows, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()
  
conf0%>%
  pivot_longer(spec_windows, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf0%>%
  pivot_longer(jinex_window, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf0%>%
  pivot_longer(acc_windows, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

```


#plot sens, spec, accuracy, j_index



```{r}
threshold = 2.5
```

```{r}
predresult <- result %>%
  select(-.model_desc, -.conf_lo, -.conf_hi, -.key) %>%
  rename(model = .model_id, value = .value, date= .index)%>%
  relocate(model, value, .after = client)%>%
  pivot_wider(names_from = model, values_from =value)%>%
  rename(actual = "NA", rf = "1", rf_tuned = "2", svm_rbd = "3", knn = "4", xgboost = "5" )%>%
  drop_na()%>%
  mutate(obs =  ifelse(actual > threshold  , TRUE, FALSE))%>%
  mutate(rf_pred=  ifelse(rf > threshold ,TRUE, FALSE))%>%
  mutate(rftuned_pred=  ifelse(rf_tuned > threshold ,TRUE, FALSE))%>%
  mutate(svm_pred=  ifelse(svm_rbd > threshold ,TRUE, FALSE))%>%
  mutate(knn_pred=  ifelse(knn > threshold ,TRUE, FALSE))%>%
  mutate(xgboost_pred=  ifelse(xgboost > threshold ,TRUE, FALSE))%>%
  mutate(obs = as.factor(obs), rf_pred = as.factor(rf_pred), rftuned_pred = as.factor(rftuned_pred), 
         svm_pred = as.factor(svm_pred),knn_pred=as.factor(knn_pred), xgboost_pred = as.factor(xgboost_pred))

conf<-predresult%>%
  group_by(date) %>%
  summarize(sens_rf = sens_vec(obs, rf_pred), 
            sens_rftuned = sens_vec(obs, rftuned_pred),
            sens_svm = sens_vec(obs, svm_pred),
            sens_knn = sens_vec(obs, knn_pred),
            sens_xgboost = sens_vec(obs, xgboost_pred),
            spec_rf = spec_vec(obs, rf_pred), 
            spec_rftuned = spec_vec(obs, rftuned_pred),
            spec_svm = spec_vec(obs, svm_pred),
            spec_knn = spec_vec(obs, knn_pred),
            spec_xgboost = spec_vec(obs, xgboost_pred),
            jindex_rf = j_index_vec(obs, rf_pred),
            jindex_rftuned = j_index_vec(obs, rftuned_pred),
            jindex_svm = j_index_vec(obs, svm_pred),
            jindex_knn = j_index_vec(obs, knn_pred),
            jindex_xgboost = j_index_vec(obs, xgboost_pred),
            acc_rf = accuracy_vec(obs, rf_pred), 
            acc_rftuned = accuracy_vec(obs, rftuned_pred),
            acc_svm = accuracy_vec(obs, svm_pred),
            acc_knn = accuracy_vec(obs, knn_pred),
            acc_xgboost = accuracy_vec(obs, xgboost_pred))

conf%>%
  pivot_longer(sens_rf:sens_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()
  
conf%>%
  pivot_longer(spec_rf:spec_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf%>%
  pivot_longer(jindex_rf:jindex_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf%>%
  pivot_longer(acc_rf:acc_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

```

```{r}
predresult1 <- result1 %>%
  select(-.model_desc, -.conf_lo, -.conf_hi, -.key) %>%
  rename(model = .model_id, value = .value, date= .index)%>%
  relocate(model, value, .after = client)%>%
  pivot_wider(names_from = model, values_from =value)%>%
  rename(actual = "NA", rf = "1", rf_tuned = "2", svm_rbd = "3", knn = "4", xgboost = "5" )%>%
  drop_na()%>%
  mutate(obs =  ifelse(actual > threshold , TRUE, FALSE))%>%
  mutate(rf_pred=  ifelse(rf > threshold ,TRUE, FALSE))%>%
  mutate(rftuned_pred=  ifelse(rf_tuned > threshold ,TRUE, FALSE))%>%
  mutate(svm_pred=  ifelse(svm_rbd > threshold ,TRUE, FALSE))%>%
  mutate(knn_pred=  ifelse(knn > threshold ,TRUE, FALSE))%>%
  mutate(xgboost_pred=  ifelse(xgboost > threshold ,TRUE, FALSE))%>%
  mutate(obs = as.factor(obs), rf_pred = as.factor(rf_pred), rftuned_pred = as.factor(rftuned_pred), 
         svm_pred = as.factor(svm_pred),knn_pred=as.factor(knn_pred), xgboost_pred = as.factor(xgboost_pred))

conf1<-predresult1%>%
  group_by(date) %>%
  summarize(sens_rf = sens_vec(obs, rf_pred), sens_rftuned = sens_vec(obs, rftuned_pred),
            sens_svm = sens_vec(obs, svm_pred),sens_knn = sens_vec(obs, knn_pred),
            sens_xgboost = sens_vec(obs, xgboost_pred),
            spec_rf = spec_vec(obs, rf_pred), spec_rftuned = spec_vec(obs, rftuned_pred),
            spec_svm = spec_vec(obs, svm_pred),spec_knn = spec_vec(obs, knn_pred),
            spec_xgboost = spec_vec(obs, xgboost_pred),
            jindex_rf = j_index_vec(obs, rf_pred), jindex_rftuned = j_index_vec(obs, rftuned_pred),
            jindex_svm = j_index_vec(obs, svm_pred),jindex_knn = j_index_vec(obs, knn_pred),
            jindex_xgboost = j_index_vec(obs, xgboost_pred),
            acc_rf = accuracy_vec(obs, rf_pred), acc_rftuned = accuracy_vec(obs, rftuned_pred),
            acc_svm = accuracy_vec(obs, svm_pred),acc_knn = accuracy_vec(obs, knn_pred),
            acc_xgboost = accuracy_vec(obs, xgboost_pred))

conf1%>%
  pivot_longer(sens_rf:sens_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()
  
conf1%>%
  pivot_longer(spec_rf:spec_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf1%>%
  pivot_longer(jindex_rf:jindex_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

conf1%>%
  pivot_longer(acc_rf:acc_xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()

```


What to calculate claim
```{r}
claim <- read_feather("process.feather")%>%
  select(date, client, claims, expected, shrinkage,volume)
```
```{r}
predclaim <-
predresult %>%
  inner_join(claim, by = c("date", "client"))%>%
  mutate(rf = rf/shrinkage *(expected /52.18),
         rf_tuned = rf_tuned/shrinkage*(expected /52.18),
         svm_rbd = svm_rbd /shrinkage*(expected /52.18),
         knn = knn/shrinkage*(expected /52.18),
         xgboost= xgboost/shrinkage*(expected / 52.18))%>%
  select(date, client, claims,expected, rf, rf_tuned,svm_rbd,knn,xgboost)

```
```{r}
predclaim%>%
  group_by(date)%>%
  summarise(expected = sum(expected)/52.18,
        claims = sum(claims),
          rf =sum(rf) ,
         rf_tuned = sum(rf_tuned),
         svm_rbd = sum(svm_rbd) ,
         knn = sum(knn),
         xgboost= sum(xgboost))%>%
  pivot_longer(expected:xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()
```

```{r}
predclaim1 <-
predresult1 %>%
  inner_join(claim, by = c("date", "client"))%>%
  mutate(rf = rf/shrinkage *(expected /52.18),
         rf_tuned = rf_tuned/shrinkage*(expected /52.18),
         svm_rbd = svm_rbd /shrinkage*(expected /52.18),
         knn = knn/shrinkage*(expected /52.18),
         xgboost= xgboost/shrinkage*(expected / 52.18))%>%
  select(date, client, claims,expected, rf, rf_tuned,svm_rbd,knn,xgboost)

```
```{r}
predclaim1%>%
  group_by(date)%>%
  summarise(expected = sum(expected)/52.18,
        claims = sum(claims),
          rf =sum(rf) ,
         rf_tuned = sum(rf_tuned),
         svm_rbd = sum(svm_rbd) ,
         knn = sum(knn),
         xgboost= sum(xgboost))%>%
  pivot_longer(expected:xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = date, y = value, color = metric)) + geom_line()
```
```{r}
predclaim%>%
  group_by(client)%>%
  
  summarise(expected = sum(expected)/52.18,
        claims = sum(claims),
          rf =sum(rf) ,
         rf_tuned = sum(rf_tuned),
         svm_rbd = sum(svm_rbd) ,
         knn = sum(knn),
         xgboost= sum(xgboost))%>%
  pivot_longer(claims:xgboost, names_to = "metric", values_to = "value")%>%
  ggplot(aes(x = client, y = value, color = metric)) + geom_line()
```

```