```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 6, dpi = 300, fig.path = "figures/time-"
)
```

# Time-based models
Here we will look at what happens when time-dependent data is added. At the moment, we only use deaths per day per zipcode

## Data wrangling
This is going to be mostly the same as `baseline_models.Rmd`. Refer to that file for more details.
Note that here we're not extracting the AEs.
```{r}
library(tidyverse)
library(tidymodels)
library(feather)
# library(arrow)
library(magrittr)
library(skimr)
library(lubridate)
library(timetk)
library(modeltime)
library(glue)
library(slider)
per <- read_feather("data/simulation_data/all_persons.feather")

clients <-
  per %>%
  group_by(client) %>%
  summarize(
    zip3 = first(zip3),
    size = n(),
    volume = sum(FaceAmt),
    avg_qx = mean(qx),
    avg_age = mean(Age),
    per_male = sum(Sex == "Male") / size,
    per_blue_collar = sum(collar == "blue") / size,
    expected = sum(qx * FaceAmt)
  )

zip_data <-
  read_feather("data/data.feather") %>%
  mutate(
    density = POP / AREALAND,
    AREALAND = NULL,
    AREA = NULL,
    HU = NULL,
    vaccinated = NULL,
    per_lib = NULL,
    per_green = NULL,
    per_other = NULL,
    per_rep = NULL,
    unempl_2020 = NULL,
    deaths_covid = NULL,
    deaths_all = NULL
  ) %>%
  rename(
    unemp = unempl_2019,
    hes_uns = hes_unsure,
    str_hes = strong_hes,
    income = Median_Household_Income_2019
  )

clients %<>%
  inner_join(zip_data, by = "zip3") %>%
  drop_na()
```

Now we have to be careful...

We take daily covid deaths and make them weekly.
Also, look at weekly deaths instead of deaths to date
```{r}
deaths <-
  read_feather("data/deaths_zip3.feather") %>%
  mutate(date = ceiling_date(date, unit = "week")) %>%
  group_by(zip3, date) %>%
  summarize(totdeaths = max(deaths)) %>%
  mutate(zip_deaths = totdeaths - lag(totdeaths, default = 0)) %>%
  select(-totdeaths) %>%
  ungroup()
```

Let's see what it looks like in Atlanta and LA. We normalize by population.
```{r atl_la_covid}
deaths %>%
  left_join(zip_data, by = "zip3") %>%
  filter(zip3 %in% c("303", "900")) %>%
  ggplot(aes(x = date, color = zip3)) +
  geom_line(aes(y = zip_deaths / POP))
```

One thing to note, there is no death data for zipcodes 202, 204, 753, 772 (which is fine, since there is no zip_data for those either).

Next we look at the deaths only
```{r}
per %<>%
  drop_na() %>%
  select(-month)
```

2019 didn't have 53 weeks, so the deaths on week 53 in 2019 will be changed to death in week 1 of 2020.

Now back to the 53 week business
```{r}
wk53 <-
  per %>%
  filter(year == 2019, week == 53) %>%
  mutate(year = 2020, week = 1)
per %<>%
  rows_update(wk53, by = c("client", "participant"))
```

The tibble `per` contains all the people that die and their deathweek. Everyone should die only once now! Let's convert the week and year into the last date of that week.

```{r}
per %<>%
  nest_by(week, year) %>%
  mutate(date = ceiling_date(ymd(glue("{year}-01-01")) + weeks(week - 1), unit = "week")) %>%
  ungroup() %>%
  select(-week, -year) %>%
  unnest(cols = c(data))
```



Next, we need some kind of a rolling count for AE. Looks like the package `slider` might help.
I want actual claims per week for each client.
We note that there are 4 clients that won't have any deaths
```{r}
no_deaths <-
  clients %>%
  anti_join(per, by = "client") %>%
  select(client) %>%
  mutate(date = ceiling_date(ymd("2019-01-01"), unit = "week"), claims = 0)
```

We compute face amount per week for each client. This number is 0 if the client has no deaths that week.
```{r}
weekly_claims <-
  per %>%
  group_by(date, client) %>%
  summarize(claims = sum(FaceAmt), .groups = "drop") %>%
  bind_rows(no_deaths) %>%
  complete(date, client, fill = list(claims = 0))
```

We now have 65,369 rows, which is 131 weeks * 499 clients.

Let's merge everything.
```{r}
weekly_data <-
  clients %>%
  left_join(weekly_claims, by = c("client")) %>%
  relocate(date) %>%
  relocate(claims, .after = zip3) %>%
  left_join(deaths, by = c("date", "zip3")) %>%
  relocate(zip_deaths, .after = claims) %>%
  mutate(zip_deaths = replace_na(zip_deaths, 0), ae = claims / (expected / 52.18))
```

We add a rolling AE number. We will smooth the actual claims number of each week by taking a weighted average of actual claims in the 13 weeks prior. The weights come from a Gaussian distribution...
The function `sliding_smoother` takes a vector and outputs a vector of smoothed values.

The below picture show what the smoother does for claims of client 7. We also show a 3 month mean AE
```{r}
smoother <- function(x) { weighted.mean(x, dnorm(seq(-1, 0, length.out = length(x)), sd = 0.33)) }
sliding_smoother <-
  slidify(smoother, .period = 13, .align = "right")

sliding_mean <-
  slidify(mean, .period = 13, .align = "right")


weekly_data %>%
  filter(client == 7) %>%
  ggplot(aes(x = date)) +
  # geom_line(aes(y = sliding_mean(claims)), color = "red") +
  # geom_line(aes(y = smooth_vec(claims, period = 13)), color = "blue") +
  geom_line(aes(y = sliding_smoother(ae)), color = "magenta") +
  geom_line(aes(y = sliding_mean(ae)), color = "red") +
  geom_line(aes(y = ae), linetype = 2)
```

We add a column called `smoothed_ae`. We may classify clients based on this number.
```{r}
weekly_data <-
  weekly_data %>%
  group_by(client) %>%
  mutate(smoothed_ae = sliding_smoother(ae), .before = size) %>%
  drop_na()
```

We plot the average AE for each week. This is pretty crazy...
```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(avg_ae = mean(smoothed_ae), sd = sd(smoothed_ae)) %>%
  ggplot(aes(x = date, y = avg_ae)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed")
  # geom_errorbar(aes(ymin = avg_ae - sd, ymax = avg_ae + sd))
```

Weekly normalized deaths in the zipcodes of our clients... I had hoped that this curve looked the same as the AE curve, but I guess it doesn't.
```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(avg_deaths = mean(zip_deaths / POP)) %>%
  ggplot(aes(x = date, y = avg_deaths)) + geom_line()
```

Let compute the 25th percentile AE of each week. We will pick the threshold for `adverse` based on this.
```{r}
weekly_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(
      `12.5` = quantile(smoothed_ae, 0.125),
      `25` = quantile(smoothed_ae, 0.25),
      `50` = quantile(smoothed_ae, 0.50)
  ) %>%
  pivot_longer(-date, names_to = "pth", values_to = "smoothed_ae") %>%
  ggplot(aes(x = date, y = smoothed_ae, color = pth)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed")
```

Maybe there is a better number to look at than smoothed ae. Let's shrink smoothed ae based on the log(Volume * average qx). This gives us some kind of a measure of client size and mortality.
```{r}
client_shrinkage <-
  weekly_data %>%
  summarize(dep_var = first(volume * avg_qx)) %>%
  mutate(shrinkage = rescale(log(dep_var), to = c(0.3, 1)))

ggplot(client_shrinkage, aes(x = dep_var)) + geom_density() + scale_x_log10()

processed_data <-
  weekly_data %>%
  left_join(client_shrinkage, by = "client") %>%
  ungroup() %>%
  mutate(shrunk_ae = smoothed_ae * shrinkage, .after = smoothed_ae)

processed_data %>%
  group_by(date) %>%
  summarize(avg_ae = mean(shrunk_ae)) %>%
  ggplot(aes(date, avg_ae)) + geom_line()


processed_data %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(
      `12.5` = quantile(shrunk_ae, 0.125),
      `25` = quantile(shrunk_ae, 0.25),
      `50` = quantile(shrunk_ae, 0.50)
  ) %>%
  pivot_longer(-date, names_to = "pth", values_to = "value") %>%
  ggplot(aes(x = date, y = value, color = pth)) +
  geom_line() +
  geom_hline(yintercept = 2.5, linetype = "dashed")
```



We will choose "smoothed, shrunk 3 month AE" > 2.5 as "Adverse".
```{r}
processed_data <-
  processed_data %>%
  ungroup() %>%
  mutate(class = factor(if_else(shrunk_ae > 2.5, "Adverse", "Not adverse"), levels = c("Adverse", "Not adverse")), .after = shrunk_ae)
```


How many clients are adverse each week?
```{r}
processed_data %>%
  group_by(date) %>%
  summarize(prop_adverse = sum(class == "Adverse") / n()) %>%
  ggplot(aes(x = date, y = prop_adverse)) + geom_line()
```



## Common data
```{r}
library(tidyverse)
library(tidymodels)
library(modeltime)
library(timetk)
library(probably)
library(themis)
library(feather)
library(magrittr)
library(skimr)
library(vip)
library(dplyr)
library(lubridate)
```

Get weeklydata from time.Rmd
```{r}
clients<-processed_data %>%
  mutate(year = year(date))%>%
  filter(year!=2019)
```

split
```{r}
#split
  split
  set.seed = 1234
  splits <-  clients %>% time_series_split(initial = "6 months", assess = "6 months", cumulative = TRUE)
  train = training(splits)
  test = testing(splits)
  #index <- unique(train$client)
  #val <- sliding_index(training(splits), index)
  
  #crossval <- vfold_cv(training(splits), strata = zip3)
```
#window forest
It is not good 
```{r}
model_spec <- window_reg(
        window_size     = "6 months",
        id = "client"
    ) %>%
    # Extra parameters passed as: set_engine(...)
    set_engine(
        engine          = "window_function",
        #window_function = median,
        na.rm           = TRUE,
        window_function = ~ tail(.x, 40),
    )

# Fit Spec
model_fit <- model_spec %>%
    fit(smoothed_ae ~ ., data = train)
model_fit
predict(model_fit, test)


model_tbl <- modeltime_table(
    model_fit)

model_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) %>%
    group_by(client) %>%
    filter(client == c( "1", "5", "7"))%>%
    plot_modeltime_forecast(
        .facet_ncol  = 3,
        .interactive = FALSE,
        .title = "Forecast Plot",
        .line_alpha = 0.6,
        .line_size = 1,
        .y_intercept = 3.0
    )
calib_tbl <- model_tbl %>%
    modeltime_calibrate(
      new_data = testing(splits), 
      id       = "client"
    )
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = FALSE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = TRUE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)

```

#Compare only one predictor(date) with all predictors(as before), the latter has better results.
#XGboost have the best result among all 7 model.

We now gather our recipes and models.
#recipe
```{r}
rec_obj <-
    recipe(smoothed_ae ~ ., data = training(splits)) %>%
    step_mutate(client = droplevels(client)) %>%
    step_timeseries_signature(date) %>%
    step_rm(date) %>%
    #step_rm(year,month,day)%>%
    step_rm(zip3)%>%
    #step_rm(adverse)%>%
    step_zv(all_predictors()) %>%
    step_dummy(all_nominal_predictors(), one_hot = TRUE)%>%
    step_normalize(all_predictors(), -all_nominal())

summary(prep(rec_obj))
bake(prep(rec_obj),training(splits))
```
```{r}
recipes <- list(rec_obj)
```

#engine
```{r}
forest_spec <-
  rand_forest(trees = 1000) %>%
  #set_mode("classification") %>%
  set_mode("regression")%>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
tuned_forest_spec <-
  rand_forest(trees = 1000, mtry = 12, min_n = 21) %>%
  #set_mode("classification") %>%
  set_mode("regression")%>%
  set_engine("ranger", num.threads = 8, importance = "impurity", seed = 123)
# Samara's models
sln_spec <-
  mlp() %>%
  set_engine("nnet") %>%
  #set_mode("classification")
  set_mode("regression")
svm_rbf_spec <-
  svm_rbf() %>%
  set_engine("kernlab") %>%
  #set_mode("classification")
  set_mode("regression")
svm_poly_spec <-
  svm_poly() %>%
  set_engine("kernlab") %>%
  #set_mode("classification")
  set_mode("regression")
knn_spec <-
  nearest_neighbor() %>%
  set_engine("kknn") %>%
  #set_mode("classification")
  set_mode("regression")
xgboost_spec <-
  boost_tree() %>%
  set_engine("xgboost") %>%
  #set_mode("classification")
  set_mode("regression")
```

# workflow 
```{r eval= FALSE}
#cannot fit
wflw_neural <- workflow() %>%
    add_model(
        sln_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))
```
```{r eval= FALSE}
#take a lot time()
wflw_svmpoly <- workflow() %>%
    add_model(
        svm_poly_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))
```

```{r}
wflw_rf <- workflow() %>%
    add_model(
        forest_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_tunedrf <- workflow() %>%
    add_model(
        tuned_forest_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_svmrbf <- workflow() %>%
    add_model(
        svm_rbf_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))

wflw_knnspec <- workflow() %>%
    add_model(
        knn_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))  

wflw_xgboost <- workflow() %>%
    add_model(
        xgboost_spec
    ) %>%
    add_recipe(rec_obj) %>%
    fit(data = training(splits))  
```
Create a Modeltime Table
```{r}
model_tbl <- modeltime_table(
    wflw_rf,
    wflw_tunedrf,
    #wflw_neural,
    wflw_svmrbf,
    #wflw_svmpoly,
    wflw_knnspec,
    wflw_xgboost
)
model_tbl
```

#Calibrate by client
```{r}
calib_tbl <- model_tbl %>%
    modeltime_calibrate(
      new_data = testing(splits), 
      id       = "client"
    )

calib_tbl
```

Measure Accuracy on validation data
```{r}
#global error
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = FALSE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
```{r}
#local error for each client
calib_tbl %>% 
    modeltime_accuracy(acc_by_id = TRUE) %>% 
    table_modeltime_accuracy(.interactive = FALSE)
```
validate the  Test Data
conf_by_id : produce confidence interval estimates by an ID feature.
#all results
```{r}
result <- calib_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = bind_rows(training(splits), testing(splits)),
        conf_by_id  = TRUE
    ) 
result
```
```{r}
result %>%
    group_by(client) %>%
    filter(client == c( "1", "5", "7","10", "61","100"))%>%
    plot_modeltime_forecast(
        .facet_ncol  = 3,
        .interactive = FALSE,
        .title = "Forecast Plot",
        .line_alpha = 0.6,
        .line_size = 1,
        .y_intercept = 3.0
    )
```

```{r}
residuals_tbl <- calib_tbl  %>%
    modeltime_calibrate(new_data = testing(splits)) %>%
    modeltime_residuals()
```
```{r}
residuals_tbl %>%
    group_by(.model_id) %>%
    #filter(.model_id == 5)%>%
    plot_modeltime_residuals(
        .type = "timeplot",
        .interactive = FALSE,
    )
```

```{r}
choose_one_client<-testing(splits)%>% filter(client == "5")
residuals_forclient <- calib_tbl  %>%
    modeltime_calibrate(new_data = choose_one_client)%>%
    modeltime_residuals()
```
```{r}
residuals_forclient %>%
    group_by(.model_id) %>%
    #filter(.model_id == 5)%>%
    plot_modeltime_residuals(
        .type = "timeplot",
        .interactive = FALSE
    )
residuals_forclient%>%
  filter(.model_id == 5)%>%
  ggplot( aes(x= .index)) +                    # basic graphical object
  geom_line(aes(y=.residuals),color="red") +# first layer
  geom_line(aes(y=.actual),color ="green") +
  geom_line(aes(y=.prediction),color ="yellow")+
  scale_fill_discrete(labels = c("residual", "actual", "prediction"))

residuals_forclient%>%
  filter(.model_id == 3)%>%
  ggplot( aes(x= .index)) +                    # basic graphical object
  geom_line(aes(y=.residuals),color="red") +# first layer
  geom_line(aes(y=.actual),color ="green") +
  geom_line(aes(y=.prediction),color ="yellow")+
  scale_fill_discrete(labels = c("residual", "actual", "prediction"))
```

```{r}
adversepred <- residuals_tbl%>%
  mutate(obs =  ifelse(.actual > 3.0, "AE>3", "AE<3"))%>%
  mutate(pred=  ifelse(.prediction > 3.0,"AE>3", "AE<3"))%>%
  select(.model_id  ,.actual,.prediction,obs,pred)
```

