# Some map visualizations

## Boring stuff

We use the sf library to handle geometries. We will also ignore AK and HI (to keep the plots nice).
```{r}
library(tidyverse)
library(sf)
library(feather)
library(ggnewscale)

geom <- st_read("zip_geom_no_ak_hi.shp")
data <- read_feather("../data/data.feather")
clients <- read_feather("../data/simulation_data/all_persons.feather")
```
## Actual plots!!
### Client plots
Zip 3 coverage
```{r}
geom %>%
  ggplot() + geom_sf(aes(fill = zip3)) + guides(fill = "none")
```

Where are our companies located?
```{r}
clients %>%
  group_by(zip3, client) %>%
  summarize() %>%
  count(sort = TRUE) %>%
  right_join(geom) %>%
  ggplot(aes(geometry = geometry, fill = n)) + geom_sf(size = 0.1) +
  scale_fill_fermenter(breaks = c(1, 2, 5, 10, 30), direction = 1, palette = "PuRd", na.value = "white") +
  theme_void() +
  labs(fill = "# of clients") +
  theme(legend.position = "bottom")
ggsave("images/location.png", dpi = 300, width = 9, height = 5)
```

### Zipcode plots
Population of zipcodes
```{r}
data %>%
  inner_join(geom) %>%
  ggplot(aes(geometry = geometry, fill = POP)) + geom_sf(size = 0.05) +
  scale_fill_fermenter(direction = 1, palette = "BuGn", na.value = "white")
```

This merged dataset will be useful to draw a bunch of maps
```{r}
merged <-
  clients %>%
  select(zip3) %>%
  unique() %>%
  mutate(has_client = TRUE) %>%
  right_join(data, by = "zip3") %>%
  replace_na(list(has_client = FALSE)) %>%
  inner_join(geom, by = "zip3")
```

Poverty in our zip3s VS others
```{r}
merged %>%
  ggplot(aes(geometry = geometry)) +
  geom_sf() +
  geom_sf(aes(fill = PCTPOVALL_2019), data = . %>% filter(has_client == TRUE)) +
  scale_fill_distiller(direction = 1) +
  labs(fill = "% poverty, clients") +
  theme(legend.position = "bottom") +
  new_scale_fill() +
  geom_sf(aes(fill = PCTPOVALL_2019), data = . %>% filter(has_client == FALSE)) +
  scale_fill_distiller(palette = "BuGn", direction = 1) +
  labs(fill = "% povery, others") +
  theme(legend.position = "bottom")
```

Our companies tend to located in more populous areas in the US
```{r}
merged %>%
  filter(has_client) %>%
  mutate(gp = "Clients") %>%
  rbind(merged %>% mutate(gp = "All US")) %>%
  ggplot(aes(y = POP, x = gp)) +
  geom_boxplot() +
  ylab("Population in Zipcode") +
  theme(axis.title.x = element_blank())
ggsave("images/population_box.png")
```

Also more wealthy
```{r}
merged %>%
  filter(has_client) %>%
  mutate(gp = "Clients") %>%
  rbind(merged %>% mutate(gp = "All US")) %>%
  ggplot(aes(y = Median_Household_Income_2019, x = gp)) +
  geom_boxplot() +
  ylab("Median household income") +
  theme(axis.title.x = element_blank())
```

Also more dense areas
```{r}
merged %>%
  filter(has_client) %>%
  mutate(gp = "Clients") %>%
  rbind(merged %>% mutate(gp = "All US")) %>%
  ggplot(aes(y = POP / AREALAND, x = gp)) +
  geom_boxplot() +
  scale_y_log10() +
  ylab("Population density") +
  theme(axis.title.x = element_blank())
```

## Misc plots
Filled column plot for AE in 2019, 2020, 2021
```{r}
clients %>%
  group_by(zip3, client) %>%
  summarize(expected = sum(qx * FaceAmt),
            actual2019 = sum(FaceAmt[year == 2019], na.rm = TRUE),
            actual2020 = sum(FaceAmt[year == 2020], na.rm = TRUE),
            actual2021 = sum(FaceAmt[year == 2021], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(AE2019 = actual2019 / expected,
         AE2020 = actual2020 / expected,
         AE2021 = actual2021 / (7 / 12 * expected)) %>%
  pivot_longer(AE2019:AE2021, names_to = "year", names_prefix = "AE", values_to = "AE") %>%
  mutate(adverse = as.factor(if_else(AE > 1, "Yes, AE > 1", "No, AE < 1"))) %>%
  select(year, adverse) %>%
  mutate(adverse = fct_reorder(adverse, desc(adverse))) %>%
  ggplot(aes(x = year, fill = adverse)) + geom_bar() +
  labs(x = "Year", y = "count", fill = "Adverse death event?")
ggsave("images/adverse.png")
```

```{r}
merged %>%
  filter(has_client) %>%
  mutate(gp = "Clients") %>%
  rbind(merged %>% mutate(gp = "All US")) %>%
  select(bachelor, `Estimated hesitant`, PCTPOVALL_2019, Unemployment_rate_2019, gp) %>%
  rename(`Percent with a bachelor` = bachelor, `Percent vaccine hesitant` = `Estimated hesitant`, `Poverty percentage` = PCTPOVALL_2019, `Unemployment rate` = Unemployment_rate_2019) %>%
  pivot_longer(-gp, names_to = "name", values_to = "value") %>%
  ggplot(aes(x = value, fill = gp)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ name, scales = "free") +
  labs(fill = "Groups")

ggsave("images/density.png", width = 16, height = 9)
```

