```{r}
  library(tidyverse)
  library(feather)
```

Read data (county,FIPS)


```{r}
population <- read_csv("Population_Estimates.csv")
vaccine <- read_csv("COVID-19_Vaccinations_in_the_United_States_County_data.gov.csv")
#death: from 01/01/2020 to 07/03/2021 total count
death <- read_csv("Provisional_COVID-19_Death_Counts_in_the_United_States_by_County.csv")
#education level percentage 15-19
education <- read_csv("Education_Estimates.csv")
poverty <- read_csv("Poverty_Estimates.csv")
unemployment <- read_csv("Unemployment_Estimates.csv")
hesitancy <- read_csv("Vaccine_Hesitancy_for_COVID-19__County_and_local_estimates.csv")
election <- read_csv("countypres_2000-2020.csv")%>%
  filter(year == 2020)
```

Read data (county name)
```{r}
lifeexpect <- read_csv("U.S._Life_Expectancy_at_Birth_by_State_and_Census_Tract_-_2010-2015.csv")
```


rename
```{r}
population <- population %>%
              rename(FIPS = FIPStxt,
                     Population = POP_ESTIMATE_2019,
                     R_Birth = R_birth_2019,
                     R_death = R_death_2019)

education <- education %>%
  rename(FIPS = `FIPS Code`,
         `less than high school` = `Percent of adults with less than a high school diploma, 2015-19`,
         `high school` = `Percent of adults with a high school diploma only, 2015-19` ,
         `college` = `Percent of adults completing some college or associate's degree, 2015-19`,
         `bachelor` = `Percent of adults with a bachelor's degree or higher, 2015-19`)

hesitancy <- hesitancy %>%
  rename(FIPS = `FIPS Code`)

vaccine <- vaccine %>%
  rename(seriescomplete = Series_Complete_Yes,
         State = Recip_State)

death <- death %>%
  rename(FIPS = `FIPS County Code`)

poverty <- poverty %>%
  rename(FIPS = FIPStxt,
         State = Stabr)

unemployment <- unemployment %>%
  rename(FIPS = FIPS_Code)

election <-election%>%
  rename(FIPS = county_fips)
```

delete useless column
```{r}
population <- population %>%
  mutate(
         R_NATURAL_INC_2019 = NULL,
         R_INTERNATIONAL_MIG_2019 = NULL,
         R_DOMESTIC_MIG_2019 = NULL,
         R_NET_MIG_2019 = NULL,
         X11 = NULL,
         State = NULL,
         Area_Name = NULL)

unemployment <- unemployment %>%
  mutate(
         Area_name = NULL,
         Area_name = NULL,
         Med_HH_Income_Percent_of_State_Total_2019 = NULL,
         State = NULL)

poverty <- poverty %>%
  mutate(
         Area_name = NULL,
         State = NULL)

death <- death %>%
  mutate(
         `Date as of` = NULL,
         `Start Date` = NULL,
         `End Date` = NULL,
         `County name` = NULL,
         `Urban Rural Code` = NULL,
         Footnote = NULL,
         State = NULL)

hesitancy <- hesitancy %>%
  mutate(
         `County Name` = NULL,
         State = NULL,
         `SVI Category` = NULL,
         `CVAC Level Of Concern` = NULL,
         `Percent Hispanic` = NULL,
         `Percent non-Hispanic American Indian/Alaska Native` = NULL,
         `Percent non-Hispanic Asian` = NULL,
         `Percent non-Hispanic Black` = NULL,
         `Percent non-Hispanic Native Hawaiian/Pacific Islander` = NULL,
         `Percent non-Hispanic White` = NULL,
         `Geographical Point` = NULL,
         `State Code` = NULL,
         `County Boundary` = NULL,
         `State Boundary` = NULL)

education <- education %>%
  mutate(
         `Area name` = NULL,
         State = NULL)

vaccine <- vaccine %>%
  mutate(
         Recip_County = NULL,
         State = NULL)

```

Make numeric FIPS into a character columns, pad with zeros
```{r}
education <- education %>%
  mutate(FIPS = str_pad(as.character(FIPS), 5, pad = "0"))

death <- death %>%
  mutate(FIPS = str_pad(as.character(FIPS), 5, pad = "0"))

hesitancy <- hesitancy %>%
  mutate(FIPS = str_pad(as.character(FIPS), 5, pad = "0"))

education <- education %>%
  mutate(FIPS = str_pad(as.character(FIPS), 5, pad = "0"))

election<-election%>%
  select(FIPS, party, candidatevotes,totalvotes)%>%
  group_by(FIPS,party)%>%
  summarise(candidatevotes=sum(candidatevotes),totalvotes= totalvotes)%>%
  distinct()%>%
  pivot_wider(names_from = party,values_from =candidatevotes)%>%
  summarise(per_dem = DEMOCRAT/totalvotes,per_rep=REPUBLICAN/totalvotes,per_other=OTHER/totalvotes,per_green=GREEN/totalvotes,per_lib=LIBERTARIAN/totalvotes) %>%
  mutate(FIPS = str_pad(as.character(FIPS), 5, pad = "0"))
```

Combine together, education has the most counties. There will be lots of NANs!!
```{r}
data <- education %>%
  left_join(population, by = "FIPS") %>%
  left_join(unemployment, by = "FIPS") %>%
  left_join(poverty, by = "FIPS") %>%
  left_join(death, by = "FIPS") %>%
  left_join(election, by = "FIPS") %>%
  left_join(hesitancy, by = "FIPS")
```

Change FIPS to zip3
```{r}
# Load the relationshop tibble
zip3_rel <- read_feather("zip3_rel.feather")
```

```{r}
cleandata <- data %>%
  mutate(state = str_sub(FIPS, 1, 2), county = str_sub(FIPS, 3, 5)) %>%
  select(-`FIPS`) %>%
  relocate(state, county)
```

```{r}
head(cleandata)
```

Clean up and aggregate by zip3. We take the weighted average of variables. Every variable will be weighted by population except household income, which will be weighted by number of households. We also remove Puerto Rico and Virgin Islands. N/A values will be ignored in our weighting.
```{r}
modifydata <- cleandata %>%
  inner_join(zip3_rel, by = c("state", "county")) %>%
  group_by(zip3) %>%
  relocate(Median_Household_Income_2019, .before = zip3) %>%
  mutate(Population = NULL) %>%
  summarize(
            across(`less than high school`:`Percent adults fully vaccinated against COVID-19 (as of 6/10/21).y`, ~ sum(.x * POPPT / sum(POPPT, na.rm = TRUE), na.rm = TRUE)),
            Median_Household_Income_2019 = sum(Median_Household_Income_2019 * HUPT / sum(HUPT, na.rm = TRUE), na.rm = TRUE),
            POP = sum(POPPT),
            HU = sum(HUPT),
            AREA = sum(AREAPT),
            AREALAND = sum(AREALANDPT)) %>%
  filter(!zip3 %in% c("006", "007", "009"))

```


```{r}
head(modifydata)
```

```{r}
write_feather(modifydata, "data.feather")
```
